SAT Competition 2020\tnoteref{title}

### 2021-06-17 paper/main.tex
results section

The traditional Hack track (est. 2009 as \solver{Minisat} Hack track) was organized as a sub-track of the Main track for hacks of \solver{Glucose 3}\textcolor{red}{Add reference}. 
In the past, several advances in SAT solving required only small modifications of an established solver to achieve a considerable contribution. 
Hack tracks encourage participation of such small modifications. 
The limit for being considered a ``hack'' was---somewhat arbitrarily\footnote{The specific threshold for edit distance is not central here; the idea is essentially to only allow relatively small changes to the Glucose
code base, i.e., ``quick hacks'' to Glucose.}---set to 1000 non-space character edit distance from the sources of \solver{Glucose 3}. 
Unfortunately, in 2020 there were not enough participants in this sub-track and so we do not report on it in the results section.

@@

In the Incremental Library track the underlying idea is to mimic scenarios
where a SAT solver is used as a back-end solver in a more complex tool
(typically solving a harder problem than SAT) and is called multiple times before 
the enclosing tool reaches its final state. 
``Incremental'' here refers to the idea that the individual calls to the SAT solver are not independent, but may share a common subset of the input clauses or differ in the presence of additional unit clause assumptions~\cite{Nadel:2014:Incremental,Fazekas:2019:IncrementalInprocessing}.\textcolor{red}{Should we not cite the 2004 temporal induction paper by Niklases here?}
Examples for applications of incremental SAT solving are counterexample-guided eastraction refinement (CEGAR) based approaches, e.g., for Bounded Model Checking~\cite{Strichman:2005:ARBMC}, SAT-based planning~\cite{Balyo:2017:IncrPlanning},
multi-agent path finding~\cite{DBLP:conf/ijcai/Surynek19}, and satisfiability modulo theories (SMT) solvers~\cite{Brummayer:2009:Boolector}.

@@

\subsection{Incremental Library Applications and Instances}

The Incremental Library track was introduced in SAT~Race~2015~\cite{Balyo:2015:SATRace} and also took place in SAT~Competitions 2016 and 2017.
Benchmarks for the Incremental Library track consist of \emph{benchmark applications} which implement and use the incremental SAT solver in their back-end as well as \emph{benchmark instances} which serve as input to these applications. 
For evaluating the Incremental Library track, we used six available IPASIR applications, ranging from determination of backbone variables to SAT-based planning. 
For each application, we individually selected 50 application instances as follows.

\paragraph{Backbone Variables}

Backbone variables~\cite{Monasson:1999,Janota:2015:Backbones} are variables which take the same value in all models of a given SAT instance.
The application \textsf{genipabones} incrementally determines backbone variables in a given satisfiable SAT instances using the dual rail encoding~\cite{Balyo:2015:SATRace}.
We selected $50$ of the smallest and easiest satisfiable instances from previous SAT competitions to evaluate solver performance with this application.

@@

Essential variables are defined to be those variables which have to be assigned in all partial models of a formula (as opposed to \emph{don't care}-values)~\cite{Bryant:1987:Essentials}. 
The application \textsf{genipaessentials} incrementally determines essential variables in a given satisfiable formula~\cite{Balyo:2015:SATRace}. 
We used the same $50$ easy satisfiable instances as in backbone detection to evaluate solver performance with this application.

@@

The application \textsf{genipalsp} determines longest simple paths in a graph~\cite{Balyo:2019:LSP}. 
We selected $50$ easy LSP instances for our evaluation.\footnote{\url{http://algo2.iti.kit.edu/kalp/}}

@@

The application \textsf{genipamax} solves partial MaxSAT problems by augmenting soft clauses with activation literals which are input to a cardinality constraint~\cite{Philipp:2015:PBLib}. 
The MaxSAT problem is then solved by incrementally minimizing the bound of that cardinality constraint. 
For this application, we selected $50$ instances from \texttt{MaxSAT Evaluation 2019}.\footnote{\url{https://maxsat-evaluations.github.io/2019/}}

\paragraph{Quantified Boolean Formulas (QBF)}
\textsf{Ijtihad} is a QBF solver which uses counter\-example-guided expansion to incrementally solve a given QBF instance with a SAT solver~\cite{Bloem:2018:QBFSAT}. 
Here we used $50$ instances from \texttt{QBF Evaluation 2019}.\footnote{\url{http://www.qbflib.org/qbfeval19.php}}

\paragraph{Planning (SAS+)}
We selected $50$ planning instances to evaluate incremental SAT solvers with \textsf{Pasar}, a planner which uses counter\-example-guided abstraction refinement (CEGAR)~\cite{Froleyks:2019:Pasar}.

@@

In this section we provide a high-level overview of the results of SAT Competition 2020.
%We focus on some interesting results.
Later on, we give more details about the solvers, including the key and new solving techniques
(Section~\ref{sec:winners}) and more in-depth analysis of the results (Section~\ref{sec:analysis}).

The four solvers \solver{Kissat-sat}, \solver{Relaxed-newTech}, \solver{CaDiCaL-alluip-trail}, and \solver{CMS-ccnr-lsids} performed significantly better that all other solvers submitted to the Main track (cf. Table~\ref{tab:results}).
Figure~\ref{main:all} shows the cumulative solved instances plot of the best-performing solver of the strongest
10 teams (in short, the top-10 solvers)  together with the Virtual Best Solver (VBS---see also~\ref{subsec:vbs_explained}).
The four top solvers all have a different code base, which is rare and has not been observed for many years.
Also, in recent competitions, the top solvers were much closer to each other in terms of runtime performance.
The winning solver is \solver{Kissat-sat} and the runner-up is \solver{Relaxed-newTech}. Notice that 
\solver{Relaxed-newTech} solved more instances within the 2000-second limit.
Third place, based on the PAR-2 score, went to \solver{CMS-ccnr-lsids}. It solved two fewer instances
compared to \solver{CaDiCaL-alluip-trail}, but solved various formulas much more quickly. 
Since the SAT Competition nowadays applies the PAR-2 score for ranking solvers, such cases have been observed several times.

@@

The main reason for the separation between the top-4 solvers and the others is the noticeable gap on satisfiable benchmarks, 
see Figure~\ref{main:sat}. In recent years, several techniques have been added to SAT solvers to improve
their performance on satisfiable instance. Examples of such techniques are the integration of a local search
solver and alternating between a SAT mode (infrequent restarts) and an UNSAT mode (frequent restarts
and variable-move-to-front~\cite{Ryan:2004:Thesis}). The 2020 winner of the Main SAT track is \solver{Relaxed-newTech}, followed by
\solver{Kissat-sat} and \solver{CMS-ccnr-lsids}.

@@

Overall, solvers performed much more similarly on
unsatisfiable formulas,
see Figure~\ref{main:unsat}. Only \solver{Kissat-unsat}, the winner of the Main UNSAT track, 
performed significantly better compared to all other participants. It is therefore not surprising
that the VBS is reasonably close to \solver{Kissat-unsat}. \solver{Kissat} is a new solver that has been implemented in C, while most solvers are implemented in C++. Further 
details about \solver{Kissat} are given in Section~\ref{sec:kissat}.
The solvers \solver{CaDiCaL-trail}
and \solver{f2trc-s} placed, respectively, second and third in the Main UNSAT track.

@@

The competition in the Planning Track was more tight. The winning solver \solver{CaDiCaL-alliup-trail} solved only one instance more compared to the runner up \solver{CMS-ccnr-lsids}. 
Their PAR-2 score is 
quite close as well. The third ranked solver, \solver{Kissat}, solved fewer instances, but its fast runtimes
on several instances resulted in a strong PAR-2 score. Notice that these three solvers were also strong in the Main track. 
Apparently none of the participating solvers was optimized for planning instances.

The winning solver in the Parallel track is \solver{Painless-MCOMSPS-STR32}. The number in the suffix denotes
that this solver used 32 threads on the 64 virtual cores that were available. It has been observed in earlier 
competitions that using fewer threads than the number of virtual cores can be helpful as threads compete 
for memory, which can reduce of the overall performance. The runner up is \solver{Plingeling}, while the
third place is taken by \solver{ManyGlucose-32}. Figure~\ref{fig:res-parallel} shows the performance of
all participating parallel solvers. Notice in Table~\ref{tab:results} that only the \solver{Painless-MCOMSPS-STR} solvers
and \solver{Plingeling} had a lower PAR-2 score than the winner of the Main Track (\solver{Kissat-sat}). 
This shows that it was hard to beat best-performing sequential solvers with a parallel solver on the competition instances. 
We will show later in the article (Table~\ref{tab:tuples}) that small sequential portfolios 
perform better (in terms of PAR-2 score) than the parallel solvers.

@@

>>>>

The traditional Hack track (established 
2009 as \solver{Minisat} Hack track) was organized as a sub-track of the Main track for hacks of \solver{Glucose 3}\textcolor{red}{Add reference}. 
In the past, several advances in SAT solving required only small modifications of an established solver to achieve a considerable contribution. 
Hack tracks encourage participation of such small modifications. 
The limit for being considered a ``hack'' was---somewhat arbitrarily\footnote{The specific threshold for edit distance is not central here; the idea is essentially to only allow relatively small changes to the Glucose
code base, i.e., ``quick hacks'' to Glucose.}---set to 1000 non-space character edit distance from the sources of \solver{Glucose 3}. 
Unfortunately, in 2020 there were not enough participants in this sub-track and so we do not report on it in the results section.

@@

The Incremental Library track was first introduced in SAT~Race~2015~\cite{Balyo:2015:SATRace} and also took place in SAT~Competitions 2016 and 2017.
In the Incremental Library track the underlying idea is to mimic scenarios
where a SAT solver is used as a back-end solver in a more complex tool
(typically solving a harder problem than SAT) and is called multiple times before 
the enclosing tool reaches its final state. 
``Incremental'' here refers to the idea that the individual calls to the SAT solver are not independent, but may share a common subset of the input clauses or differ in the presence of additional unit clause assumptions~\cite{Nadel:2014:Incremental,Fazekas:2019:IncrementalInprocessing}.\textcolor{red}{Should we not cite the 2004 temporal induction paper by Niklases here?}
Examples for applications of incremental SAT solving are counterexample-guided eastraction refinement (CEGAR) based approaches, e.g., for Bounded Model Checking~\cite{Strichman:2005:ARBMC}, SAT-based planning~\cite{Balyo:2017:IncrPlanning},
multi-agent path finding~\cite{DBLP:conf/ijcai/Surynek19}, and satisfiability modulo theories (SMT) solvers~\cite{Brummayer:2009:Boolector}.

@@

\subsection{Incremental Library Track Benchmarks}

Benchmarks for the Incremental Library track consist of \emph{benchmark applications} which implement and use the incremental SAT solver in their back-end as well as \emph{benchmark instances} which serve as input to these applications. 
For evaluating solvers participating in the 
the Incremental Library track, we used six available IPASIR applications.
For each of the six applications, we individually selected 50 application instances as follows.

\paragraph{Backbone Computation}

Backbone variables~\cite{Monasson:1999,Janota:2015:Backbones} are variables which take the same value in all models of a given SAT instance.
The application \textsf{genipabones} incrementally determines backbone variables in a given satisfiable SAT instances using the so-called 
dual rail encoding~\cite{Balyo:2015:SATRace}.
We selected $50$ of the smallest and easiest satisfiable instances from previous SAT competitions to evaluate solver performance with this application.

@@

Variables which have to be assigned in all partial models of a formula as \emph{essential} (as opposed to \emph{don't care}-values)~\cite{Bryant:1987:Essentials}. 
The application \textsf{genipaessentials} incrementally determines essential variables in a given satisfiable formula~\cite{Balyo:2015:SATRace}. 
For this applicaiton, we used the same $50$  satisfiable instances as for  backbone computation.

@@

The application \textsf{genipalsp} determines longest simple paths in a graph~\cite{Balyo:2019:LSP}. 
We selected $50$ LSP instances for our evaluation.\footnote{\url{http://algo2.iti.kit.edu/kalp/}}

@@

The application \textsf{genipamax} solves partial MaxSAT problems by augmenting soft clauses with relaxation (or blocking) variables
which are input to a cardinality constraint~\cite{Philipp:2015:PBLib}. 
The MaxSAT problem is then solved by incrementally minimizing the bound of the cardinality constraint. 
For this application, we selected $50$ instances from \texttt{MaxSAT Evaluation 2019}.\footnote{\url{https://maxsat-evaluations.github.io/2019/}}

\paragraph{Quantified Boolean Formulas (QBF)}
\textsf{Ijtihad} is a QBF solver which uses counterexample-guided expansion to incrementally solve a given QBF instance with a SAT solver~\cite{Bloem:2018:QBFSAT}. 
Here we used $50$ instances from \texttt{QBF Evaluation 2019}.\footnote{\url{http://www.qbflib.org/qbfeval19.php}}

\paragraph{Planning (SAS+)}
We selected $50$ planning instances to evaluate incremental SAT solvers with \textsf{Pasar}, a planner which uses counterexample-guided abstraction refinement (CEGAR)~\cite{Froleyks:2019:Pasar}.

@@

In this section we provide a high-level overview of the results of SAT Competition 2020.
%We focus on some interesting results.
Later on, we will provide overview of some of the key and new solving techniques implemented in best-performing solvers
(Section~\ref{sec:winners}) and as well as more in-depth analysis of the competition results (Section~\ref{sec:analysis}).

\section{Main Track}

Starting with the Main track, 
Figure~\ref{main:all} shows the cumulative solved instances plot of the best-performing solver of the strongest
10 teams (in short, the top-10 solvers)  together with the Virtual Best Solver (VBS---see also~\ref{subsec:vbs_explained}).
The best-performing  solver overall on the combination of satisfiable and unsatisfiable instances
is \solver{Kissat-sat} and the runner-up is \solver{Relaxed-newTech}. Notice that
\solver{Relaxed-newTech} solved more instances within the 2000-second limit.
Third place, based on the PAR-2 score, went to \solver{CMS-ccnr-lsids}. It solved two  instances less than
\solver{CaDiCaL-alluip-trail}, but on the other hand solved various formulas more quickly.
Similar observations have been made also in earlier recent SAT competitions where 
solvers were ranked based on the PAR-2 score.
Furthermore, we observe more differences in overall runtime performance among the top solvers than what has been observed
in the recent past competitions.

The four solvers \solver{Kissat-sat}, \solver{Relaxed-newTech}, \solver{CaDiCaL-alluip-trail}, and \solver{CMS-ccnr-lsids} 
performed significantly better that all other solvers submitted to the Main track (cf. Table~\ref{tab:results}).
A very interesting observation is that these four top solvers all have a different code base. This has not been observed for many years; more typically
many of the best-performing solver have been based on same code bases.

The majority of the overall performance differences
 between the top-4 solvers and the other solvers 
is due to performance differences on satisfiable instances;
see Figure~\ref{main:sat}. Indeed, in recent years, several techniques have been added to SAT solvers to improve
their performance on satisfiable instance. Examples of such techniques are the integration of a local search
solver and alternating between a SAT mode (infrequent restarts) and an UNSAT mode (frequent restarts
and variable-move-to-front~\cite{Ryan:2004:Thesis}). The best-performing solver in  Main SAT track on satisfiable instances is \solver{Relaxed-newTech}, followed by
\solver{Kissat-sat} and \solver{CMS-ccnr-lsids}.

@@

@@

Overall, solvers performed much more similarly on
unsatisfiable instances than on satisfiable instances;
see Figure~\ref{main:unsat} for the runtime performance on unsatisfiable instances. 
Only \solver{Kissat-unsat}, the winner of the Main UNSAT track, 
performed significantly than all other participating solvers. 
It is therefore not surprising
that the VBS is reasonably close to \solver{Kissat-unsat}. %\solver{Kissat} is a new solver that has been implemented in C, while most solvers are implemented in C++. Further 
%details about \solver{Kissat} are given in Section~\ref{sec:kissat}.
The solvers \solver{CaDiCaL-trail}
and \solver{f2trc-s} placed, respectively, second and third in the Main UNSAT track.

@@

\subsection{Planning Track}

The competition in the Planning Track was more tight. 
The best solver \solver{CaDiCaL-alliup-trail} solved only one instance more than  the runner up \solver{CMS-ccnr-lsids}. 
The PAR-2 scores of these two solvers were
quite similar as well. The third ranked solver, \solver{Kissat}, solved fewer instances, but its fast runtimes
on several instances resulted in a strong PAR-2 score. Notice that these three solvers were also strong in the Main track. 
It should be noted that, somewhat disappointingly,  none of the participating solvers were actually optimized for planning instances.

\textcolor{red}{Hmm we don't have any result tables or anything??}

\subsection{Parallel Track}

Turning to the Parallel track, Figure~\ref{fig:res-parallel} shows the performance of
all participating parallel solvers. 
The best solver in the Parallel track is \solver{Painless-MCOMSPS-STR32}. Interestingly,
this solver used 32 threads on the 64 virtual cores that were available. In fact, it has been observed in already 
recent earlier 
SAT competitions that using fewer threads than the number of available virtual cores can be helpful; as threads compete 
for memory, using all virtual cores may be detrimental to  overall performance. The runner up is \solver{Plingeling}, while the
third place goes to \solver{ManyGlucose-32}. 
Interestingly, one can observe from Table~\ref{tab:results} that only the \solver{Painless-MCOMSPS-STR*} solvers
and \solver{Plingeling} had a lower PAR-2 score than the winner of the Main Track (\solver{Kissat-sat}). 
It appears still to be challenging to beat the best-performing sequential solvers with
with a parallel solver. (In fact, as we will show later on, sequential portfolios including only a few
Main track solvers show  better performance than all of the parallel solvers.)

@@

\subsection{Cloud Track}
